Kaggle-in-class Data Challenges Can Boost Student Learning

### 2019-09-23 paper-kaggle.Rmd
confoudning

MAST90083 is titled Computational Statistics and Data Mining, is designed for postgraduate level, for students with math, statistics, information technology or actuarial backgrounds. It covers modelling both continuous (regression) and categorical (classification) response variables. The 63 students were randomized into one of two kaggle competitions, one focused on regression (R) and the other classification (C). This setup mimics randomized control trials, which is the 'gold standard, in experiment design (@Shelley09CH1, ch. 1).
Students built prediction models and made submissions individually for 16 days, and then were allowed to form groups to compete for another 7 days.

The reason for this strategy was first to motivate each of the students to think about modelling and be actively engaged in the competitions through individual submission. The lecturer allowed to create groups towards the end of the competition to illustrate the advantages of group work and mixing models. Another motivation for this stratagem was the university policy, requiring a strategy to assign students individually in group assignments.

0ne of the 63 students didn't took part in the competition. Another student didn't sit the exam. Therefore, the analysed sample size reduced to 61.

ETC2420/5242, titled Statistical Thinking, covers regression, and has a mix of undergraduate and postgraduate students. Only the 34 postgraduate (5242) students were required to participate in the kaggle competition, and competed in the regression (R) challenge. This was run independently from the University of AB competition. The 145 undergraduate (2420) students were used as controls for examining performance of the postgraduate students. Although, it may be surprising at first glance, the undergraduate students provide a reasonable control for the graduate students. The class is taught to both cohorts simultaneously. The entry requirements to the Bachelors of Commerce at Monash is high, and these students have strong mathematics backgrounds. In the years prior to this experiment, the undergraduate scores on the final exam are indistinguishable from those of the graduate students. The competition ran for one month. Students formed their own teams of 2-4 members to compete. 
<!--
Several undergraduates also chose to compete individually. The material on regression methods, particularly the computational methods needed to successfully predict housing prices was new to both groups of students. 
-->
<!--
ETC3250, called Business Analytics, is an undergraduate course focusing on data mining. All students participated in a kaggle competition on the classification challenge. Because this group had no comparison group, it was difficult to assess performance. 
-->

@@

This paper has described the setup and results of an experiment to examine the effectiveness of data competitions on student learning, using Kaggle InClass as the vehicle for conducting the competition. The experiment was conducted in the classroom settings as part of the normal teaching of the courses, which imposes limitations on the design. However, with both MAST90083 and ETC2420/5242, either a randomized assignment of students to two topic groups, or a control group was possible, which enabled comparing performance.

@@

This work is one of a few quantitative analyses of data competition influences on students' performance. More evidence needs to be collected from other STEM courses to demonstrate consistent positive influence. Moreover, future investigation is required to understand the influence of the different aspects of data competition implementation on the magnitude of the performance improvement. For example, the competition duration, availability and accessibility of additional material, requirement of writing a final report or giving a short oral presentation are elements worth investigating. In this work we couldn't provide statistically significant evidence for positive influence of the data competition on the engagement and the interest. This should be addressed in a future work. Finally, it will be interesting to investigate if some sub-group of students benefits more from proposed competitions, e.g. undergraduate vs graduate students or female vs male students.

>>>>

Computational Statistics and Data Mining (MAST90083) is designed for postgraduate level  students with math, statistics, information technology or actuarial backgrounds. It covers modelling both continuous (regression) and categorical (classification) response variables. The 63 students were randomized into one of two kaggle competitions, one focused on regression (R) and the other classification (C). This setup mimics randomized control trials, which is the 'gold standard, in experiment design (@Shelley09CH1, ch. 1).
Students built prediction models and made submissions individually for 16 days, and then were allowed to form groups to compete for another 7 days.

The reason for this strategy was first to motivate each of the students to think about modelling and be actively engaged in the competitions through individual submission. The lecturer allowed participants to create groups towards the end of the competition to illustrate the advantages of group work and mixing models. Another motivation for this stratagem was the university policy, requiring a strategy to assign students individually in group assignments.

One of the 63 students elected not to take part in the competition, and another student didn't sit the exam, producing a final sample size of 61.

Statistical Thinking (ETC2420/5242), covers regression, but not classification, and has a mix of undergraduate and postgraduate students. Only the 34 postgraduate (5242) students were required to participate in the kaggle competition, and competed in the regression (R) challenge. This was run independently from the MAST90083 competition. The 145 undergraduate (2420) students were used as controls for examining performance of the postgraduate students. Although, it may be surprising at first glance, the undergraduate students provide a reasonable control for the graduate students. The class is taught to both cohorts simultaneously. The entry requirements to the Bachelors of Commerce at Monash is high, and these students have strong mathematics backgrounds. In the years prior to this experiment, the undergraduate scores on the final exam are indistinguishable from those of the graduate students. The competition ran for one month. Students formed their own teams of 2-4 members to compete. 
<!--
Several undergraduates also chose to compete individually. The material on regression methods, particularly the computational methods needed to successfully predict housing prices was new to both groups of students. 
-->
<!--
ETC3250, called Business Analytics, is an undergraduate course focusing on data mining. All students participated in a kaggle competition on the classification challenge. Because this group had no comparison group, it was difficult to assess performance. 
-->

@@

This paper has described the setup and results of an experiment to examine the effectiveness of data competitions on student learning, using Kaggle InClass as the vehicle for conducting the competition. The experiment was conducted in the classroom settings as part of the normal teaching of the courses, which imposes limitations on the design. However, with both MAST90083 and ETC2420/5242, either a randomized assignment of students to two topic groups, or a control group, was possible, which enabled comparing performance.

@@

This work is one of a few quantitative analyses of data competition influences on students' performance. More evidence needs to be collected from other STEM courses to explore consistent positive influence. Moreover, future investigation is required to understand the influence of the different aspects of data competition implementation on the magnitude of the performance improvement. For example, the competition duration, availability and accessibility of additional material, requirement of writing a final report or giving a short oral presentation are elements worth investigating. Prior and post testing of students might provide a better experimental design. <!--In this work we couldn't provide statistically significant evidence for positive influence of the data competition on the engagement and the interest. This should be addressed in a future work.Finally, it will be interesting to investigate if some sub-groups of students benefit more from competitions, e.g. undergraduate vs graduate students or female vs male students.-->
